Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-10-20 draft.tex
[draft] Add some more questions

As before, we assume the PDF $P$ can be decomposed into $N$ parts with angular observables $\lbrace S_k\rbrace$,
\begin{equation}
    P \equiv P(\vec\nu,\vec\theta | \lbrace S_k \rbrace)\,.
\end{equation}
We denote the estimator of an
observable $S_k$ as $\est{S_k}$, which is affected by the detector acceptance. We assume now that there is a linear relation
between the true angular observables $\lbrace S_k^\text{true}\rbrace$, and the estimators $\lbrace \est{S_k}\rbrace$, \fred{We don't need the $^{\text true}$.}
\begin{equation}
    \label{eq:linear-assumption}
    \est{S_i} = \sum_{j = 1}^N M_{i j} S_j^\text{true}\,.
\end{equation}
This assumption is based on a smooth angular acceptance function. While this will not be true for the laboratory frame,
smoothness may be recovered in a boosted frame, e.g. the rest frame of the decaying particle.
In that case, the matrix $M_{ij}$ can be fully determined through the following steps.
\begin{enumerate}
    \item For a fixed $j$, let $S_j^\text{true} = 1$, and $S_k^\text{true} = 0\,\forall k\neq j$.
    \item Draw simulated events $\lbrace x_t^{(k)}\rbrace = \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ from the PDF $P_j$,
        \begin{equation}
            P_j \equiv P(\vec\nu,\vec\theta | \lbrace S_j = 1, S_{k\neq j} = 0\rbrace)\,.
        \end{equation}
    \item Use the true events as input for the MC simulation of the detector, in order to obtain the detector
        events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,\vec\theta^{(k)}_d)\rbrace$.
    \item Determine the uncorrected observables $\lbrace \est{S_k} \rbrace$ based on the detector
        events $\lbrace x^{(k)}_d\rbrace$. For all intents and purposes, automation of this step is only feasible
        with to the method of moments, since it does not suffer form convergence problems.
    \item By virtue of \refeq{linear-assumption} and step 1, the $j$-th column of $M$ has been determined via
        \begin{equation}
            M_{ij} = \est{S_i}\,.
        \end{equation}
    \item Repeat the above steps for all values of $j$ to determine all columns of $M$.
\end{enumerate}

@@

We emphasize that this method hinges on the assumption in \refeq{linear-assumption}\danny{which is well motivated:
the PDF is a linear combination of the basis functions $f_i$, and the moments are therefore a linear
functional of said basis functions. Nevertheless,}
we thus suggest to explicitly
test the linearity. One way of testing involves repeating the previous procedure
with $S_j^\text{true} = 1/z$, $z > 0$. If the linearity assumption holds, one should find
\begin{equation}
    M_{ij}(z) = z \est{S_i}\,,
\end{equation}
for a fixed value of $j$.

@@

The PDF for the decay $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ --- up to and including P-wave contributions --- has been calculated
for the most general basis of dimension-six $b\to s$ operators. It reads, expressed in terms of the angular observables $\lbrace J_i\rbrace$ \cite{Blake:2012mb,Bobeth:2012vn}
\begin{equation}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3) = \frac{1}{\rmdx{\Gamma}/\rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}} = \frac{3}{8\pi} \sum_k \frac{J_{k}(q^2)}{\rmdx{\Gamma} /\, \rmdx{q^2}} f_k(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{equation}
where $\theta_1 \equiv \theta_\ell$ is the dilepton helicity angle, and $\theta_2 \equiv \theta_{K}$ is the $\bar{K}\pi$ helicity angle, and $\theta_3 \equiv \phi$ is the azimuthal angle.
The $q^2$-differential decay width reads
\begin{equation}
    \frac{\rmdx{\Gamma}}{\rmdx{q^2}} = \frac{\big(3 J_{1c} - J_{2c}\big) + 2\big(3J_{1s} - J_{2s}\big)}{3}\,.
\end{equation}
The angular functions\footnote{%
    We consistently use the label "$i$" to denote angular function and observables that arise
    from interference between S-wave and P-wave amplitudes. Our convention is related to the convention
    of \cite{Bobeth:2012vn} through $f_{1i} \leftrightarrow f_{1sc}$, $f_{2i} \leftrightarrow f_{2sc}$.
} read \cite{Blake:2012mb,Bobeth:2012vn}
\begin{equation}
\begin{aligned}
    f_{1s} & = \sin^2\theta_2 &
    f_{1c} & = \cos^2\theta_2 &
    f_{1i} & = \cos\theta_2\\
    f_{2s} & = \sin^2\theta_2 \cos 2\theta_1 &
    f_{2c} & = \cos^2\theta_2 \cos 2\theta_1 &
    f_{2i} & = \cos\theta_2 \cos 2\theta_1\\
    f_{3}  & = \sin^2\theta_2\sin^2\theta_1 \cos 2\theta_3 &
    f_{4}  & = \sin 2\theta_2 \sin 2\theta_1 \cos\theta_3 &
    f_{4i} & = \sin\theta_2 \sin 2\theta_1 \cos\theta_3\\
    f_{5}  & = \sin 2\theta_2 \sin \theta_1 \cos\theta_3 & & &
    f_{5i} & = \sin\theta_2 \sin \theta_1 \cos\theta_3\\
    f_{6s} & = \sin^2\theta_2 \cos\theta_1 &
    f_{6c} & = \cos^2\theta_2 \cos\theta_1\\
    f_{7}  & = \sin 2\theta_2 \sin \theta_1 \sin\theta_3 & & &
    f_{7i} & = \sin\theta_2 \sin \theta_1 \sin\theta_3\\
    f_{8}  & = \sin 2\theta_2 \sin 2\theta_1 \sin\theta_3 & & &
    f_{8i} & = \sin\theta_2 \sin 2\theta_1 \sin\theta_3\\
    f_{9}  & = \sin^2\theta_2\sin^2\theta_1 \sin 2\theta_3\\
\end{aligned}
\end{equation}
We indicate contributions from the interference of S waves and P waves by an index \emph{i}. In terms of the angles $\theta_1$ and $\theta_2$, instead of their cosines, the PDF reads
\begin{equation}
    P(q^2, \theta_1, \theta_2, \theta_3) = \frac{3}{8\pi} \sum_k S_k(q^2) f_k(\cos\theta_1, \cos\theta_2, \theta_3) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $S_k(q^2) \equiv J_k(q^2) / (\rmdx{\Gamma} /\, \rmdx{q^2})$.\\

>>>>

As before, we assume the PDF $P$ can be decomposed into $N$ parts with angular observables $\lbrace S_k\rbrace$,
\begin{equation}
    P \equiv P(\vec\nu,\vec\theta | \lbrace S_k \rbrace)\,.
\end{equation}
We denote the estimator of an
observable $S_k$ as $\est{S_k}$, which is affected by the detector acceptance. We assume now that there is a linear relation
between the true angular observables $\lbrace S_k^\text{true}\rbrace$, and the estimators $\lbrace \est{S_k}\rbrace$, \fred{We don't need the $^{\text true}$.}
\begin{equation}
    \label{eq:linear-assumption}
    \est{S_i} = \sum_{j = 1}^N M_{i j} S_j^\text{true}\,.
\end{equation}
This assumption is based on a smooth angular acceptance
function. \fred{Check the LHCb \href{http://inspirehep.net/record/1287929}{paper} on $B \to K \ell \ell$: their
  acceptance (fig. 1) is not smooth due to a cut!} \fred{I don't understand. Need a derivation of why this hinges on smoothness} While this will not be true for
the laboratory frame, smoothness may be recovered in a boosted frame,
e.g. the rest frame of the decaying particle.  In that case, the
matrix $M_{ij}$ can be fully determined through the following steps.
\begin{enumerate}
    \item For fixed $j$, let $S_j^\text{true} = 1$, and $S_k^\text{true} = 0\,\forall k\neq j$.
    \item Draw simulated events $\lbrace x_t^{(k)}\rbrace =
      \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ \fred{don't use $k$ as superscript and subscript} from the PDF $P_j$,
      \begin{equation}
        \label{eq:single-comp}
            P_j \equiv P(\vec\nu,\vec\theta | \lbrace S_j = 1, S_{k\neq j} = 0\rbrace)\,.
        \end{equation}
    \item Use the true events as input for the MC simulation of the detector in order to obtain the detector
        events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,\vec\theta^{(k)}_d)\rbrace$.
    \item Determine the uncorrected observables $\lbrace \est{S_k} \rbrace$ based on the detector
        events $\lbrace x^{(k)}_d\rbrace$. For all intents and purposes, automation of this step is only feasible
        with to the method of moments, since it does not suffer form convergence problems.
    \item By virtue of \refeq{linear-assumption} and step 1, the $j$-th column of $M$ has been determined via
        \begin{equation}
            M_{ij} = \est{S_i}\,.
        \end{equation}
    \item Repeat the above steps for all values of $j$ to determine all columns of $M$.
\end{enumerate}

@@

We emphasize that this method hinges on the assumption in \refeq{linear-assumption}\danny{which is well motivated:
the PDF is a linear combination of the basis functions $f_i$, and the moments are therefore a linear
functional of said basis functions. Nevertheless,}
we thus suggest to explicitly
test the linearity. One way of testing involves repeating the previous procedure
with $S_j^\text{true} = 1/z$, $z > 0$. If the linearity assumption holds, one should find
\begin{equation}
    M_{ij}(z) = z \est{S_i}\,,
\end{equation}
for a fixed value of $j$. \fred{This is useless. \refeq{single-comp}
  says: draw from a single component $f_j$. This $f_j$ defines the
  distribution, and whatever factor you put in front of it, you get
  the same distribution of events.}

@@

The PDF for the decay $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ --- up to and including P-wave contributions --- has been calculated
for the most general basis of dimension-six $b\to s$ operators. It reads, expressed in terms of the angular observables $\lbrace J_i\rbrace$ \cite{Blake:2012mb,Bobeth:2012vn}
\begin{equation}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3) = \frac{1}{\rmdx{\Gamma}/\rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}} = \frac{3}{8\pi} \sum_k \frac{J_{k}(q^2)}{\rmdx{\Gamma} /\, \rmdx{q^2}} f_k(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{equation}
where $\theta_1 \equiv \theta_\ell$ is the dilepton helicity angle, and $\theta_2 \equiv \theta_{K}$ is the $\bar{K}\pi$ helicity angle, and $\theta_3 \equiv \phi$ is the azimuthal angle.
The $q^2$-differential decay width reads
\begin{equation}
    \frac{\rmdx{\Gamma}}{\rmdx{q^2}} = \frac{\big(3 J_{1c} - J_{2c}\big) + 2\big(3J_{1s} - J_{2s}\big)}{3}\,.
\end{equation}
The angular functions\footnote{%
    We consistently use the label "$i$" to denote angular function and observables that arise
    from interference between S-wave and P-wave amplitudes. Our convention is related to the convention
    of \cite{Bobeth:2012vn} through $f_{1i} \leftrightarrow f_{1sc}$, $f_{2i} \leftrightarrow f_{2sc}$.
} read \cite{Blake:2012mb,Bobeth:2012vn}
\begin{equation}
\begin{aligned}
    f_{1s} & = \sin^2\theta_2 &
    f_{1c} & = \cos^2\theta_2 &
    f_{1i} & = \cos\theta_2\\
    f_{2s} & = \sin^2\theta_2 \cos 2\theta_1 &
    f_{2c} & = \cos^2\theta_2 \cos 2\theta_1 &
    f_{2i} & = \cos\theta_2 \cos 2\theta_1\\
    f_{3}  & = \sin^2\theta_2\sin^2\theta_1 \cos 2\theta_3 &
    f_{4}  & = \sin 2\theta_2 \sin 2\theta_1 \cos\theta_3 &
    f_{4i} & = \sin\theta_2 \sin 2\theta_1 \cos\theta_3\\
    f_{5}  & = \sin 2\theta_2 \sin \theta_1 \cos\theta_3 & & &
    f_{5i} & = \sin\theta_2 \sin \theta_1 \cos\theta_3\\
    f_{6s} & = \sin^2\theta_2 \cos\theta_1 &
    f_{6c} & = \cos^2\theta_2 \cos\theta_1\\
    f_{7}  & = \sin 2\theta_2 \sin \theta_1 \sin\theta_3 & & &
    f_{7i} & = \sin\theta_2 \sin \theta_1 \sin\theta_3\\
    f_{8}  & = \sin 2\theta_2 \sin 2\theta_1 \sin\theta_3 & & &
    f_{8i} & = \sin\theta_2 \sin 2\theta_1 \sin\theta_3\\
    f_{9}  & = \sin^2\theta_2\sin^2\theta_1 \sin 2\theta_3\\
\end{aligned}
\end{equation}
In terms of the angles $\theta_1$ and $\theta_2$, instead of their cosines, the PDF reads
\begin{equation}
    P(q^2, \theta_1, \theta_2, \theta_3) = \frac{3}{8\pi} \sum_k S_k(q^2) f_k(\cos\theta_1, \cos\theta_2, \theta_3) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $S_k(q^2) \equiv J_k(q^2) / (\rmdx{\Gamma} /\, \rmdx{q^2})$.\\
